### 2019-07-17 manuscript/methods.tex
Fixed all todos and revised discussion issues

The Supplementary Information contains citations for the numerous original data sources. % FIXME (and say here how many refs?) %B: no need for an additional counting and updating headache?
The Dryad archive contains the data and tree files used for analyses. % FIXME
% \url{https://github.com/roszenil/solploidy}  E: we should use dryad instead, because github repos are not guaranteed to last

@@

The simplest model (M16) classifies each species as either SI diploid $(ID)$, SC diploid $(CD)$, or SC polyploid $(CP)$; recall that SI polyploids do not occur.
Each of these states may again be associated with different rates of speciation and extinction, and the allowable transitions are loss of SI within the diploid state (from $ID$ to $CD$), loss of SI via polyploidization (from $ID$ to $CP$), and polyploidization while SC (from $CD$ to $CP$).
% The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
% E: Can we call it something other than q_IC?  It's a bit confusing to have the same parameter mean different things in different models.

@@

Second, we consider not only transitions between states but also diversification within each state.
State-dependent diversification can change the relative contributions of the two pathways.
For example, if the net diversification rate is small for $CD$, the two-step pathway will contribute relatively less.
We therefore include the difference between speciation and extinction along the diagonal elements of the propagation matrix.
As before, matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
In this case, however, these are not probabilities because diversification changes the number of lineages as time passes. %B: FIXME, we'll hear complaints again about "this," "these," and ... "their." (below)
We can still use ratios, however, to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population. % R- here a reference would be nice  E: Does Jake have a favorite ref? Otherwise, I guess a basic ecology textbook.
%
% E & R todo ?: for the case with net div, write out the propagation matrix and equations for pathway contributions
%
% E: I don't think we need this section now that diploidization is in each of the sections above.  But I'm saving this text in case we want to put some of it in Results.
%--------------------------------------------------
% \subsubsection{Diploidization as an exploratory hypothesis}
%
% The models described above allow polyploidization but not its reverse.
% We explored the effects of allowing diploidization by adding a parameter $\delta$ for transitions from $P$ back to $D$ (in the $D/P$ and $D/P+A/B$ models) or from $CP$ back to $CD$ (in the $I/C$ and $I/C+A/B$ models); we did not allow transitions from $CP$ back to $ID$ because that would include regain of SI.
% Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010, glick2014, mayrose_2015, freyman_2017}, which assume only aneuploid change and polyploidization.
% Our ploidy classifications, however, when possible also incorporated information on chromosome multiplicity at the genus level and author conclusions within original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution).
% Since it is not known whether diploidization can be detected under these additional ploidy classifications, we fit models with and without diploidization in order to test whether the conclusions about diversification are sensitive to including diploidization.
% In our models, the diploidization parameter (or its absence, $\delta=0$), represents an opportunity to explore a possibly significant assumption that may not decisively affect interactions among ploidy, breeding system, and diversification.
%--------------------------------------------------

@@

Parameters for each of the 29 models were estimated using RevBayes \citep{hoehna_2016}.
Scripts for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.
We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the models.
Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.

@@

We calculated the marginal log-likelihood for each of the models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps. % (\cref{table:marginallike}) updated ref?

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{states.pdf}
\caption{
Character states used for each of the models.
Each species retained on the tree belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known. Character state abbreviations: $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
(A) The five categories and the number of species in our data for each category.
Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC ($CP$).
(B) Category groupings into states for each model type.
For example, there are 70 species that are SC with unknown ploidy.
In the $D/P$ models these are coded as $D$ and $P$ (uncertain/consistent with either state), in the $I/C$ models they are coded as $C$, and in the $ID/CD/CP$ models they are coded as $CD$ and $CP$.
In the hidden state models, all species were coded as consistent with both $A$ and $B$.
Two species, \emph{Lycium californicum} and \emph{Solanum bulbocastanum}, are simultaneously $ID$ and $CP$, and by adding them the sample adds to the total of 651 taxa used for analyses.
}
\label{figure:stateclassifications}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{alldiversificationasymhidden.pdf} 
\caption{Net diversification rates for all models without diploidization rate estimate. 
Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}
\label{figure:netdivall}
\end{figure}

@@

\begin{supptable}
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|l|r|c|c|c|c|l|}
\hline
Model                            & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M18   & M19   & M20   & Evidence                                                                                                   \\ \hline
M16. ID/CP/CD                    & -1459.11                                                          & 45.11 & 65.91 & 63.98 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M16\end{tabular}                         \\ \hline
M17. CID ID/CP/CD                & *                                                                 &       &       &       &                                                                                                            \\
M18. ID/CP/CD+A/B                & -1414                                                             &       & 20.79 & 18.87 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}               \\ \hline
\textbf{M19. ID/CP/CD +A/B asym} & \textbf{-1393.20}                                                 &       &       & -1.92 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates moderately\\ preferred over all asymmetric\end{tabular} \\ \hline
M20. ID/CP/CD +A/B all asym      & -1393.12                                                          &       &       &       &   \\ \hline                                                                                                        
\end{tabular}
\caption{Bayes factors of ploidy and breeding system withou diploidization. Results indicate that the MuHiSSE model with asymmetric hidden rates (17) is  strongly preferred over M16 and M18 and moderately preferred over the MuHiSSe with all rates asymetric (M20). *Marginal log-likelihood for M17 could not be calculated within allotted computer time.}
\label{supptable:M16M20}
\end{supptable}

@@

\begin{supptable}
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|l|c|c|c|c|}
\hline
Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
M1. D/P                                  & -1238.76                                                          & M1 vs. M6   & 65.92           & M6 (Strong)                                                          \\
\textbf{M6. D/P+$\delta$}                & \textbf{-1267.84}                                                 &             &                 &                                                                      \\ \hline
M4. D/P+A/B asym                         & -1223.28                                                          & M4 vs. M9   & 8.81            & M9 (Moderate)                                                        \\
\textbf{M9. D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
M16. ID/CD/CP                            & -1459.11                                                          & M16 vs. M21 & 4.41            & M21 (Moderate)                                                       \\
\textbf{M21. ID/CD/CP+$\delta$}          & \textbf{-1454.68}                                                 &             &                 &                                                                      \\ \hline
M19. IC/CD/CP asym                       & -1393.20                                                          & M19 vs. M24 & 6.46            & M24 (Moderate)                                                       \\
\textbf{M24. IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
\end{tabular}
\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidization are moderately preferred over models that do not include a diploidization rate}
\label{supptable:testdiploidization}
\end{supptable}

>>>>

The Supplementary Information contains citations for the numerous original data sources. % FIXME (and say here how many refs?) %B: no need for an additional counting and updating headache?
% R- There is few data that I added so I can find those references and send them to you, so it would be simply the same references that you previously had plus another 20-25 I think
The Dryad archive contains the data and tree files used for analyses. % FIXME
% \url{https://github.com/roszenil/solploidy}  E: we should use dryad instead, because github repos are not guaranteed to last
%This fixme happens after publications

@@

The simplest model (M16) classifies each species as either SI diploid $(ID)$, SC diploid $(CD)$, or SC polyploid $(CP)$; recall that SI polyploids do not occur.
Each of these states may again be associated with different rates of speciation and extinction, and the allowable transitions are loss of SI within the diploid state (from $ID$ to $CD$), loss of SI via polyploidization (from $ID$ to $CP$), and polyploidization while SC (from $CD$ to $CP$).
% The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
% E: Can we call it something other than q_IC?  It's a bit confusing to have the same parameter mean different things in different models.
% R: In the lumped models I call them something different q_0 and rho_0 because they are different parameters but I don't think we should explain.

@@

Second, we consider not only transitions between states but also diversification within each state.
State-dependent diversification can change the relative contributions of the two pathways.
For example, if the net diversification rate is small for $CD$, the two-step pathway will contribute relatively less.
We therefore include the difference between speciation and extinction along the diagonal elements of the propagation matrix.
As before, matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
The calculations of the propagation matrix are not probabilities because diversification changes the number of lineages as time passes. 
%B: FIXME, we'll hear complaints again about "this," "these," and ... "their." (below)
% R- fixed
We can still use ratios, however, to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population. % R- here a reference would be nice  E: Does Jake have a favorite ref? Otherwise, I guess a basic ecology textbook.
%
% E & R todo ?: for the case with net div, write out the propagation matrix and equations for pathway contributions
% R- couldn't solve to get the exact solutions turns out it is really hard.
%
% E: I don't think we need this section now that diploidization is in each of the sections above.  But I'm saving this text in case we want to put some of it in Results.
%--------------------------------------------------
% \subsubsection{Diploidization as an exploratory hypothesis}
%
% The models described above allow polyploidization but not its reverse.
% We explored the effects of allowing diploidization by adding a parameter $\delta$ for transitions from $P$ back to $D$ (in the $D/P$ and $D/P+A/B$ models) or from $CP$ back to $CD$ (in the $I/C$ and $I/C+A/B$ models); we did not allow transitions from $CP$ back to $ID$ because that would include regain of SI.
% Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010, glick2014, mayrose_2015, freyman_2017}, which assume only aneuploid change and polyploidization.
% Our ploidy classifications, however, when possible also incorporated information on chromosome multiplicity at the genus level and author conclusions within original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution).
% Since it is not known whether diploidization can be detected under these additional ploidy classifications, we fit models with and without diploidization in order to test whether the conclusions about diversification are sensitive to including diploidization.
% In our models, the diploidization parameter (or its absence, $\delta=0$), represents an opportunity to explore a possibly significant assumption that may not decisively affect interactions among ploidy, breeding system, and diversification.
%--------------------------------------------------

@@

Parameters for each of the 29 models were coded as graphical models and Bayesian statistical inference was performed RevBayes \citep{hoehna_2016}.
Scripts for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.- Happens after publication
We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the models.
Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.

@@

We calculated the marginal log-likelihood for each of the models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:bayesfactors}).

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{states.pdf}
\caption{
Character states used for each of the models.
Each species retained on the tree belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known. Character state abbreviations: $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
(A) The five categories and the number of species in our data for each category.
Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC ($CP$).
(B) Category groupings into states for each model type.
For example, there are 70 species that are SC with unknown ploidy.
In the ploidy only models (M1- M10) these are coded as $D$ and $P$ (uncertain/consistent with either state), in the breeding system only models (M11-M15) they are coded as $C$, and in the ploidy and breeding system models (M16-M29) they are coded as $CD$ and $CP$.
In the hidden state models, all species were coded as consistent with both $A$ and $B$.
Two species, \emph{Lycium californicum} and \emph{Solanum bulbocastanum}, are simultaneously $ID$ and $CP$, and by adding them the sample adds to the total of 651 taxa used for analyses.
}
\label{figure:stateclassifications}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{alldiversificationasymhidden.pdf} 
\caption{Net diversification rates for SSE models of focal traits with or without hidden state. 
Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}
\label{figure:netdivall}
\end{figure}

@@

\begin{supptable}
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|l|r|c|c|c|c|l|}
\hline
Model                            & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M18   & M19   & M20   & Evidence                                                                                                   \\ \hline
M16. ID/CP/CD                    & -1459.11                                                          & 45.11 & 65.91 & 63.98 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M16\end{tabular}                         \\ \hline
M17. CID ID/CP/CD                & *                                                                 &       &       &       &                                                                                                            \\
M18. ID/CP/CD+A/B                & -1414                                                             &       & 20.79 & 18.87 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}               \\ \hline
\textbf{M19. ID/CP/CD +A/B asym} & \textbf{-1393.20}                                                 &       &       & -1.92 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates moderately\\ preferred over all asymmetric\end{tabular} \\ \hline
M20. ID/CP/CD +A/B all asym      & -1393.12                                                          &       &       &       &   \\ \hline                                                                                                        
\end{tabular}
\caption{Bayes factors of ploidy and breeding system without diploidization. Results indicate that the MuHiSSE model with asymmetric hidden rates (17) is  strongly preferred over M16 and M18 and moderately preferred over the MuHiSSe with all rates asymetric (M20). *Marginal log-likelihood for M17 could not be calculated within allotted computer time.}
\label{supptable:M16M20}
\end{supptable}

@@

\begin{supptable}
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|l|c|c|c|c|}
\hline
Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
M1. D/P                                  & -1238.76                                                          & M1 vs. M6   & 65.92           & M6 (Strong)                                                          \\
\textbf{M6. D/P+$\delta$}                & \textbf{-1267.84}                                                 &             &                 &                                                                      \\ \hline
M4. D/P+A/B asym                         & -1223.28                                                          & M4 vs. M9   & 8.81            & M9 (Moderate)                                                        \\
\textbf{M9. D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
M16. ID/CD/CP                            & -1459.11                                                          & M16 vs. M21 & 4.41            & M21 (Moderate)                                                       \\
\textbf{M21. ID/CD/CP+$\delta$}          & \textbf{-1454.68}                                                 &             &                 &                                                                      \\ \hline
M19. IC/CD/CP +A/B asym                       & -1393.20                                                          & M19 vs. M24 & 6.46            & M24 (Moderate)                                                       \\
\textbf{M24. IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
\end{tabular}
\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidization are moderately preferred over models that do not include a diploidization rate}
\label{supptable:testdiploidization}
\end{supptable}
